### 2018-11-12 source/Ch4-Qweights/paper.tex
asdasda

\pg
The main result of this this section is a new adaptive, quality-based weighting scheme based on this insight. Using the Cov-Cov relationship, we can {create a new} weighting scheme by estimating the residual visibility covariance matrix in a given observation. By weighting {visibilities} so as to change their covariance matrix, one can change the shape of the noise-PSF and thus improve {the final image}: this manifests as either decreased noise or decreased calibration artefacts. Note that this weighting is applied after calibration, but before image deconvolution: applying it will therefore not only improve the residual noise in the image, and thus the sensitivity achievable with a given pipeline, but will also improve deconvolution by minimising calibration artefacts in the field: it should thus effectively remove spurious, unphysical emission from final data products. Estimating the covariance matrix is the main difficulty of our framework: we do not know the underlying covariance matrix, and the conditioning of our estimation thereof is limited by the number of measurements within each {solution interval}. As such, we have no guarantee that our estimate of the corrected visibility covariance matrix is accurate. This problem can be alleviated, for example {by estimating the covariance matrix for the antenna gains themselves}, and use {it} to build the visibility covariance matrix: this effectively improves conditioning (cf Sec. \ref{section.algorithm}).

@@

\pg
$\bm{R}$ corresponds to a matrix containing all the residual visibilities within one calibration cell $\cellindex$, i.e. for $b \in \cellindex$ where $\gvect_{\cellindex} = \const$ {It is therefore of size $n_{\mathrm{intervals}}\times n_{\cellindex}$, where $n_{\mathrm{intervals}}$ is the number of calibration intervals in the observation}. Normalising the residual visibilities by $k_b$ allows us to recover the underlying covariance matrix by multiplying the residual visibility matrix $\bm{R}$ with its Hermitian conjugate:
\begin{align}
\estC[b\in\cellindex,b'\in\cellindex']             =& \left(\bm{R}^H \bm{R}\right)[\cellindex,\cellindex']
\end{align}
Note that we have divided the noise term by the flux model $S_b$, which can be very small in some cells. As such, care must be taken not to cause the relative thermal noise contribution to explode: those cells where this would occur are dominated by thermal noise, and information on the covariance matrix cannot be recovered from them.

@@

\pg
In estimating the covariance matrix for each baseline and each calibration cell, we are severely limited by the small number of samples in each cell. One way to overcome this problem is to find estimates for the variance of \emph{antenna} gains, and use these to return to the baseline variances. In this formalism, we extend  $\cellindex$ to include all visibilities pointing at a single antenna at a given time. Let us begin by writing an expression for the gain vector, which contains the gains for all antennas and all calibration cells:
\begin{align}
\ghatvect_c =& \begin{pmatrix} \vdots \\ \hat{g}_p^{\tnuIndeix\in c} \\ \vdots \end{pmatrix}\\
\Ghatvect =& \begin{pmatrix} \hdots \quad \ghatvect_c \quad \hdots \end{pmatrix}
\end{align}
and the variance on each antenna in each calibration cell is then:
\begin{align}
               =& \Exp{\ghatvect_c \ghatvect_c^H} - \underbrace{\Exp{\ghatvect_c} \Exp{\ghatvect_c}^H}_{= \gvect_c \gvect_c^H} \label{eq.def.vector.variance}
\end{align}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-goodcal.png}}
\caption{\label{image.3c295.goodcal} Well-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. Used for comparison with the other images. {Units of colourbar are Janskys}. This image was made with data calibrated following best practice (solution intervals of 8 seconds, half the bandwidth). rms=5.87mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-noweights.png}}
\caption{\label{image.3c295.nocorr} Poorly-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. {Units of colourbar are Janskys}. This image was made with the same data as for Fig. \ref{image.3c295.goodcal}, but averaged in time and calibrated using larger gain solution intervals: 2 minutes and half the bandwidth. rms=86.4mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-cyrilweights.png}}
\caption{\label{image.3c295.lightcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, using sensitivity-optimal weighting. {Units of colourbar are Janskys}. rms=9.69mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-fullweights.png}}
\caption{\label{image.3c295.fullcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, using artefact-optimal weighting. {Units of colourbar are Janskys}. rms=15.8mJy/beam}
\end{subfigure}
\caption{\label{image.fourRealImages}Restored images of the centre of the Extended Groth Strip, as seen with an 8-hour observation using the full LOFAR array. Fig. \ref{image.3c295.goodcal} shows an image of the field made with good calibration intervals. Fig. \ref{image.3c295.nocorr} shows an image of the field made with poor calibration intervals. Fig. \ref{image.3c295.lightcorr} shows image made with the same visibilities and imaging parameters, but with the application of the sensitivity-optimal weighting scheme. Fig. \ref{image.3c295.fullcorr}, similarly, differs from Fig \ref{image.3c295.lightcorr} only in that artefact-optimal weights, rather than sensitivity-optimal weights, were used. {The histograms of pixel values in each image have 1000 flux bins ranging from -0.16 Jy to 0.16 Jy. Their ordinates are in log scale.} Pixel size is 1.5'' in all images.}
\end{figure}

\pg
Note that applying antenna-based sensitivity-optimal weighting to the badly-calibrated data (not shown here) allows us to recover the reference image with only a very small increase in rms (increased by a factor of 1.14). Further testing on complex field simulations will be required to ascertain the usefulness of artefact-optimal weighting: it is likely that it fails to correct the image fully due to the poor conditioning of the covariance matrix used here.

@@

\pg
{The pixel histograms show us that the weights do not completely mitigate the poor calibration interval choice, but certainly give a dramatic improvement over the unweighted, poorly-calibrated residuals. This is compatible with our statement that the weights give similar residuals in the image with a dramatic improvement in time at some cost in sensitivity. It is interesting to note that while Fig. \ref{image.3c295.fullcorr} looks noisier than Fig \ref{image.3c295.lightcorr}, its residual flux histogram is actually closer to that of Fig. \ref{image.3c295.goodcal}.}

\pg
As for performance, the weights used for Fig. \ref*{image.3c295.fullcorr} took 8 hours of computing time on a single core\footnote{Core type: Intel(R) Xeon(R) CPU E5-2660 0 @ 2.20GHz}, working on a 29 GB dataset, which is not particularly large for LOFAR data. Since the problem is massively parallel, this cost can be alleviated. The main bottleneck is likely due to very poor code optimization. As for the weights used for Fig. \ref*{image.3c295.antcorr}, they are computed in 1min 6s on the same single core.

@@

\pg
%In effect, we have provided a basis to justify the use of unpublished, but still commonly-used, methods like ``clipping'' statistically outlying calibration gains. 
{While our results are not a panacea for poor calibration, they show that we can not only improve images made with well-calibrated data, but also mitigate the worst effects of poorly-calibrated visibilities in otherwise well-calibrated datasets. Provided that the gain variability timescale is long enough at certain points of the observation, we can effectively get images of similar quality using both the} ``standard'' best-practice calibration interval for LOFAR survey data (calibration solution interval of 8 seconds) and a significantly larger solution interval of 2 minutes (frequency interval unchanged). {Of course, if no such stable interval exists, there will be no good intervals to upweigh, and we will be left only with equally-poor data chunks.} This means that{, in the right conditions,} net pipeline time {can be} sped up by a factor of nearly three, at a slight cost in sensitivity. {This increase will be greater than what could be achieved with existing comparable methods such as ``clipping''.} However, it does not aim to replace such methods, but rather to act as a complement to them: clipping will be better at flagging a handful of bad visibilities, while this weighting scheme will help push the general visibility statistics (and therefore the image-plane statistics) towards a better distribution.

\pg
We emphasize that the adaptive quality-based weighting schemes work because the noise-{map} describes the \emph{underlying noise distribution}, of which calibration artefacts are \emph{one single realisation}. To fully characterise the artefacts, the correlation between different pixels (i.e. off-diagonal elements of $\Cov{\muely}$) must be computed; this has not been done in this paper. Nevertheless, lesser constraints on the spatial distribution of artefacts can be found using only the diagonal elements of $\Cov{\muely}$. The weighting schemes merely seek to {minimise} this spatial distribution as much as possible: the end result is \emph{fewer artefacts}, {which can be} distributed across a \emph{much larger area}. {This} is the source of the dramatic improvement from \ref{image.3c295.nocorr} to Fig. \ref{image.3c295.lightcorr}. We have simply down-weighted those visibilities where spurious signal was introduced by the calibration solutions, and up-weighted those visibilities where such signal was lesser. Since this spurious signal is the source of calibration artefacts, downweighting the associated visibilities reduces it dramatically. {The poor improvement from Fig. \ref{image.3c295.nocorr} to \ref{image.3c295.fullcorr}  is likely due to limits in the conditioning of our estimation of the covariance matrix.}

>>>>

\pg
The main result of this section is a new adaptive, quality-based weighting scheme based on this insight. Using the Cov-Cov relationship, we can {create a new} weighting scheme by estimating the residual visibility covariance matrix in a given observation. By weighting {visibilities} so as to change their covariance matrix, one can change the shape of the noise-PSF and thus improve {the final image}: this manifests as either decreased noise or decreased calibration artefacts. Note that this weighting is applied after calibration, but before image deconvolution: applying it will therefore not only improve the residual noise in the image, and thus the sensitivity achievable with a given pipeline, but will also improve deconvolution by minimising calibration artefacts in the field: it should thus effectively remove spurious, unphysical emission from final data products. Estimating the covariance matrix is the main difficulty of our framework: we do not know the underlying covariance matrix, and the conditioning of our estimation thereof is limited by the number of measurements within each {solution interval}. As such, we have no guarantee that our estimate of the corrected visibility covariance matrix is accurate. This problem can be alleviated, for example {by estimating the covariance matrix for the antenna gains themselves}, and use {it} to build the visibility covariance matrix: this effectively improves conditioning (cf Sec. \ref{section.algorithm}).

@@

\pg
$\bm{R}$ corresponds to a matrix containing all the residual visibilities within one calibration cell $\cellindex$, i.e. for $b \in \cellindex$ where $\gvect_{\cellindex} = \const$ {It is therefore of size $n_{\mathrm{intervals}}\times n_{\cellindex}$, where $n_{\mathrm{intervals}}$ is the number of calibration intervals in the observation}. Normalising the residual visibilities by $k_b$ allows us to recover the underlying covariance matrix by multiplying the residual visibility matrix $\bm{R}$ with its Hermitian conjugate:
\begin{align}
\estC[b\in\cellindex,b'\in\cellindex']             =& \left(\bm{R}^H \bm{R}\right)[\cellindex,\cellindex']
\end{align}
Note that we have divided the noise term by the flux model $\kappa_b$, which can be very small in some cells. As such, care must be taken not to cause the relative thermal noise contribution to explode: those cells where this would occur are dominated by thermal noise, and information on the covariance matrix cannot be recovered from them.

@@

\pg
In estimating the covariance matrix for each baseline and each calibration cell, we are severely limited by the small number of samples in each cell. One way to overcome this problem is to find estimates for the variance of \emph{antenna} gains, and use these to return to the baseline variances. In this formalism, we extend  $\cellindex$ - the set of visibilities in a calibration cell -  to include all visibilities pointing at a single antenna at a given time. Let us begin by writing an expression for the gain vector, which contains the gains for all antennas and all calibration cells:
\begin{align}
\ghatvect_c =& \begin{pmatrix} \vdots \\ \hat{g}_p^{\tnuIndeix\in c} \\ \vdots \end{pmatrix}\\
\Ghatvect =& \begin{pmatrix} \hdots \quad \ghatvect_c \quad \hdots \end{pmatrix}
\end{align}
and the variance on each antenna in each calibration cell is then:
\begin{align}
               =& \Exp{\ghatvect_c \ghatvect_c^H} - \underbrace{\Exp{\ghatvect_c} \Exp{\ghatvect_c}^H}_{= \gvect_c \gvect_c^H} \label{eq.def.vector.variance}
\end{align}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-goodcal.png}}
\caption{\label{image.3c295.goodcal} Well-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. Used for comparison with the other images. {Units of colourbar are Janskys}. This image was made with data calibrated following best practice (solution intervals of 8 seconds, half the bandwidth). rms=5.87mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-noweights.png}}
\caption{\label{image.3c295.nocorr} Poorly-calibrated, unweighted restored image of the sky near the centre of the Extended Groth Strip. {Units of colourbar are Janskys}. This image was made with the same data as for Fig. \ref{image.3c295.goodcal}, but averaged in time and calibrated using larger gain solution intervals: 2 minutes and half the bandwidth. rms=86.4mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-cyrilweights.png}}
\caption{\label{image.3c295.lightcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, using sensitivity-optimal weighting. {Units of colourbar are Janskys}. rms=9.69mJy/beam}
\end{subfigure}
\hfill
\begin{subfigure}{.49\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/paperfig-fullweights.png}}
\caption{\label{image.3c295.fullcorr} Image made using the same imaging parameters and corrected visibilities as Fig. \ref{image.3c295.nocorr}, using artefact-optimal weighting. {Units of colourbar are Janskys}. rms=15.8mJy/beam}
\end{subfigure}
\caption{\label{image.fourRealImages}Restored images of the centre of the Extended Groth Strip, as seen with an 8-hour observation using the full LOFAR array. Fig. \ref{image.3c295.goodcal} shows an image of the field made with good calibration intervals. Fig. \ref{image.3c295.nocorr} shows an image of the field made with poor calibration intervals. Fig. \ref{image.3c295.lightcorr} shows image made with the same visibilities and imaging parameters, but with the application of the sensitivity-optimal weighting scheme. Fig. \ref{image.3c295.fullcorr}, similarly, differs from Fig \ref{image.3c295.lightcorr} only in that artefact-optimal weights, rather than sensitivity-optimal weights, were used. All weighting schemes applied in these figures are baseline-based rather than antenna-based. {The histograms of pixel values in each image have 1000 flux bins ranging from -0.16 Jy to 0.16 Jy. Their ordinates are in log scale.} Pixel size is 1.5'' in all images.}
\end{figure}

\pg
Note that applying antenna-based sensitivity-optimal weighting to the badly-calibrated data (shown in \cref{image.3c295.antcorr}) allows us to recover the reference image with only a very small increase in rms (increased by a factor of 1.14). Further testing on complex field simulations will be required to ascertain the usefulness of artefact-optimal weighting: it is likely that it fails to correct the image fully due to the poor conditioning of the covariance matrix used here.

@@

\pg
{The pixel histograms show us that the weights do not completely mitigate the poor calibration interval choice, but certainly give a dramatic improvement over the unweighted, poorly-calibrated residuals. This is compatible with our statement that the weights give similar residuals in the image with a dramatic improvement in time at some cost in sensitivity. It is interesting to note that while Fig. \ref{image.3c295.fullcorr} may look slightly noisier than Fig \ref{image.3c295.lightcorr}, its residual flux histogram is actually closer to that of Fig. \ref{image.3c295.goodcal}.}

\pg
%As for performance, the weights used for Fig. \ref*{image.3c295.fullcorr} took 8 hours of computing time on a single core\footnote{Core type: Intel(R) Xeon(R) CPU E5-2660 0 @ 2.20GHz}, working on a 29 GB dataset, which is not particularly large for LOFAR data. Since the problem is massively parallel, this cost can be alleviated. The main bottleneck is likely due to very poor code optimization. As for the weights used for Fig. \ref*{image.3c295.antcorr}, they are computed in 1min 6s on the same single core. }

@@

\pg
%In effect, we have provided a basis to justify the use of unpublished, but still commonly-used, methods like ``clipping'' statistically outlying calibration gains. 
{While our results are not a panacea for poor calibration, they show that we can not only improve images made with well-calibrated data, but also mitigate the worst effects of poorly-calibrated visibilities in otherwise well-calibrated datasets. Provided that the gain variability timescale is long enough at certain points of the observation, we can effectively get images of similar quality using both the} ``standard'' best-practice calibration interval for LOFAR survey data (calibration solution interval of 8 seconds) and a significantly larger solution interval of 2 minutes (frequency interval unchanged). {Of course, if no such stable interval exists, there will be no good intervals to upweigh, and we will be left only with equally-poor data chunks.} This means that{, in the right conditions,} net pipeline time {can be} sped up by a factor of nearly three (by increasing the size of the calibration solution intervals), at a slight cost in sensitivity. {This increase will be greater than what could be achieved with existing comparable methods such as ``clipping''.} However, it does not aim to replace such methods, but rather to act as a complement to them: clipping will be better at flagging a handful of bad visibilities, while this weighting scheme will help push the general visibility statistics (and therefore the image-plane statistics) towards a better distribution.

\pg
We emphasize that the adaptive quality-based weighting schemes work because the noise-{map} describes the \emph{underlying noise distribution}, of which calibration artefacts are \emph{one single realisation}. To fully characterise the artefacts, the correlation between different pixels (i.e. off-diagonal elements of $\Cov{\muely}$) must be computed; this has not been done in this thesis. Nevertheless, lesser constraints on the spatial distribution of artefacts can be found using only the diagonal elements of $\Cov{\muely}$. The weighting schemes merely seek to {minimise} this spatial distribution as much as possible: the end result is \emph{fewer artefacts}, {which can be} distributed across a \emph{much larger area}. {This} is the source of the dramatic improvement from \ref{image.3c295.nocorr} to Fig. \ref{image.3c295.lightcorr}. We have simply down-weighted those visibilities where spurious signal was introduced by the calibration solutions, and up-weighted those visibilities where such signal was lesser. Since this spurious signal is the source of calibration artefacts, downweighting the associated visibilities reduces it dramatically. {The poor improvement from Fig. \ref{image.3c295.nocorr} to \ref{image.3c295.fullcorr}  is likely due to limits in the conditioning of our estimation of the covariance matrix.}
